Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2013-12-29 paper.Rnw
Added latest FEP xval results

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: prep
% Prepares the data in a the form needed for plotting.
<<ADNI1-xval-load,include=FALSE,cache=TRUE>>=
jens_atlases = read.csv('data/jens_atlases.csv')
jens_xval_data  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data.csv.gz'))
jens_xval_mean  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data_mean.csv.gz'))
jens_xval_mean$reg_method = factor(jens_xval_mean$reg_method, c("ANTS", "ANIMAL"))
jens_xval_mean$method.mb = factor(jens_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
mb.kappa.best = mean(subset(jens_xval_mean, 
                                 reg_method   == "ANTS" & 
                                 method.mb    == "Majority Vote" &
                                 templates.mb == 19 & 
                                 atlases      == 9)$k.mb)
mb.diff.best = mean(subset(jens_xval_mean, 
                                 reg_method   == "ANTS" & 
                                 method.mb    == "Majority Vote" &
                                 templates.mb == 19 & 
                                 atlases      == 1)$k_diff)
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, url, amsmath, geometry, ctable}
\usepackage{titlesec, siunitx, graphicx}
\usepackage[round,authoryear]{natbib}
\usepackage[section]{placeins}   % keep floats in their place.
\usepackage{authblk}             % Listing Author affiliations
\usepackage{rotating}            % rotating figures
\usepackage[hypcap]{caption}
\usepackage{lineno}

%draft mode
\usepackage[colorinlistoftodos, textwidth=4cm, shadow]{todonotes}
\usepackage[displaymath, tightpage]{preview}
\setlength{\marginparwidth}{1.2in}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[RO,RE]{draft - December, 2013}

@@

\textbf{Introduction:} 
Advances in image segmentation of magnetic resonance images (MRI) have
demonstrated that \multiatlas{} approaches improve segmentation accuracy and
precision over regular atlas-based approaches.  These approaches often rely on
a large number of such manually segmented atlases (e.g.  30-80) that take
significant time and expertise to produce. We present an algorithm, \mb{} 
({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
T}emplates), for the automatic segmentation of the hippocampus that minimizes
the number of atlases needed while still achieving similar accuracy to
\multiatlas{} approaches. Thus, our method acts as an accurate \multiatlas{}
approach when using special, hard-to-define atlases that are laborious to
construct. \\
\textbf{Method:} \mb{} works by propagating atlas segmentations to a template
library, formed from a subset of target images, via transformations estimated by
nonlinear image registration. The resultant segmentations are then propagated
to each target image and fused using a label fusion method.  To explore the
effect of atlases and templates library size, registation and label
fusion method, we conduct a 10-fold Monte Carlo cross-validation of \mb{} of whole
hippocampal segmentation on 60 ADNI subjects, manually segmented using the
\prsnr{}-protocol, over a range of parameter settings. Using the best settings
found, we then conduct a leave-one-out cross-validation (LOOCV) of hippocampal
subfield segmentation using five high-resolution manually segmented atlases.
Two final experiments assess \mb{} when applied to first episode psychosis and
Alzheimer's disease populations, and \mb{} segmentations are compared with
existing automated methods (\fsl{}, \freesurfer{}, \maper{}) and biases are
explored.\\
\textbf{Results:} Using 9 atlases and 19 template images, \mb{} achieves a
mean Dice's Similarity Coefficient (DSC) of \Sexpr{round(mb.kappa.best, digits=3)} 
(to \psegs{}) over 10-folds of Monte Carlo % TODO: update cross-validation on 60 subjects, and shows
significantly lower variability in DSC than \multiatlas{} segmentation.
In a LOOCV, \mb{} reproduces the volumes of the cornu ammonis (CA) 1;
CA4/dentate % TODO: update
gyrus (DG); and strata radiatum (SR), strata lacunosum (SL), and strata
moleculare (SM) hippocampal subfields with a percent error in volume that is
at or lower than that produced by image resampling. \mb{} produces hippocampal
volumes in a first episode psychosis patient population that are highly
correlated with expert manual segmentation volumes (Pearson $r = 0.877, t =
16.244, p < 0.001$). Compared to \fsl{} and \freesurfer{}, \mb{} shows much smaller
fixed volume bias (within $250 mm^3$ on average) to semi-automated (\snt{})
segmentations available from ADNI, as well as a conservative, rather than
exaggerated, proportional volume bias.\\ \textbf{Conclusion:} We demonstrate
that \mb{} produces accurate hippocampal segmentations using only 5 atlases over
different hippocampal definitions, disease populations, and acquisition types,
as well as showing that accurate identification of the hippocampal subfields is
possible.
\end{abstract}

@@

We find that for \mb{} segmentations, similarity score increases
as atlas and template library size is increased, although with diminishing returns and
an eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}).  A maximum
similarity score of \Sexpr{round(mb.kappa.best, digits=3)} is found when using 9
atlases, 19 templates, ANTS registration, and majority vote label fusion).  The
\ants{} registration method consistently out-performs \animal{} registration
over all variable settings we tested. Additionally, by itself, using a weighted
voting strategy did not significantly improve segmentation accuracy, contrary to
the findings of \citet{Aljabar2009} using basic \multiatlas{} segmentation.
Given these findings, in the remainder of this section only results using the
\ants{} registration algorithm and majority vote fusion will be shown.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
\begin{figure}
<<fig:ADNI1-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=7, fig.height=7>>=
stderr <- function(x) sqrt(var(x)/length(x))
ggplot(subset(jens_xval_mean, (templates.mb*atlases) %% 2 == 1 & templates.mb < 21),
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases), linetype=reg_method)) + #as.factor((atlases*templates.mb) %% 2 ))) + 
  facet_grid(.~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
  stat_summary(fun.y=mean,geom='point',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
  stat_summary(fun.y=mean,
               fun.ymin=function (x) mean(x) - stderr(x),
               fun.ymax=function (x) mean(x) + stderr(x), 
               geom='errorbar',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases))), linetype='solid', width=0.5) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_y_continuous(breaks=seq(0,1,0.01)) + 
  scale_colour_hue(name="Number of Atlases") +
  scale_linetype(name="Registration Method") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean similarity (DSC)" ) + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text = element_text(size = 8))
@
  \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
  \psegs{} for 60 ADNI subjects vs. atlas and template library
  size, registration algorithm, and label fusion method. Error bars indicate
  standard error.
  \label{fig:ADNI1-xval-k-mean}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: "increase" in mean kappa over \multiatlas{}
\begin{figure}
<<fig:ADN11-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-load'>>=
ggplot(subset(jens_xval_mean, reg_method=="ANTS" & 
                method.mb=="Majority Vote" & (templates.mb*atlases) %% 2 == 1), 
       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_summary(fun.y=mean, geom="line", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Number of Atlases") +
  scale_alpha_continuous(guide = "none") + 
  scale_linetype_discrete(guide = "none") + 
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean similarity (DSC)" ) +
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{Increase in mean Dice's Similarity Coefficient of \mb{} over
  \multiatlas{} segmentations vs. atlas and template library size when using the
  \ants{} registration method, and majority-vote label fusion.
  Segmentation similarity is computed against \psegs{}.  
  Error bars indicate standard error.
  \label{fig:ADNI1-xval-k-diff}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
  #facet_grid( . ~ stat) + 
  geom_line(size=0.5) +
  geom_point(size=2) + 
  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,by=2)) + 
  xlab( "Number of Templates" ) + 
  ylab( "Variability (p)") +
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  scale_size_continuous(guide='none') + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas{} segmentation
  with respect to manual segmentation.}
  Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation
  is computed for each subject across all ten rounds of validation. 
  Shown on the y-axis is the  p-value resulting from a t-test comparing the
  distribution of variances at each parameter setting (atlas/template library
  size). 
  Only points where \mb{} mean variability is lower than \multiatlas{} are shown.
  Dashed lines indicate a p-value of 0.05 and 0.01.}
  \label{fig:ADNI1-xval-variability}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

ggplot(data=fep.volumes, aes(x=manual, y=volume, colour=method)) + 
  geom_smooth(method="lm", formula=y~x) + 
  geom_point() + 
  geom_text(aes(x=3350,y=y,label=eqn),parse=T,data=fep.lm,
      size=3,hjust=0) +
  xlab(expression("Mean manual hippocampus volume " (mm^3))) + 
  ylab(expression("Mean computed hippocampus volume "  (mm^3))) + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{Mean bilateral hippocampus volume as measured by \mb{}, \fsl{}, and
\freesurfer{} vs. manually segmented volumes from the First Episode Patients with
Schizophrenia dataset. \mb{} is run two ways: with an atlas library consistening
of the five Winterburn atlases, or with five manually segmented FEP subjects as
atlases (not included as subjects). Linear fit lines are shown, with shaded
region showing standard error. 
  \label{fig:SZ-volumes}}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tabulation of segmentation accuracies of existing methods (ADNI)
%
%
\begin{table}[!tbp]
\scriptsize
\caption{\textbf{Automated segmentation accuracy (overlap with \snt{} labels)
of the ADNI dataset.} For each method, the number of labelled atlases
used for training, the best Dice's overlap measure, the disease classes
measured, and the validation procedure are shown. Unless specified, validation
datasets are composed equally of subjects diagnosed with Alzheimer's Disease
(AD), Mild Cognitive Impairment (MCI), and Cognitively Normal (CN). LOOCV =
Leave-one-out cross-validation. 
   Some studies of automated segmentation of ADNI images are excluded because they 
   do not not provide overlap measures for the hippocampus \citep{Heckemann2011,Chupin2009}.
  \label{tab:other-methods}} 
\begin{center}
\begin{tabular}{ l  c  p{1in}  l  p{2in}  }
\hline\hline
Method & Atlases & DSC & Reference & Validation \tabularnewline
\hline
\mb{}                        &9 & \Sexpr{round(mb.kappa.best, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline % TODO: update
LEAP                         &30&0.848                 &\cite{Wolz2010}  &Segmentation of 182 subjects\tabularnewline
ACM (AdaBoost-based)         &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline
Patch-based label fusion     &16&0.883 (CN) \newline 0.838 (AD)&\cite{Coupe2011e}&LOOCV on atlases\tabularnewline
Multi-atlas                  &30&0.885                 &\cite{Lotjonen2010}&Segmentation of 60 subjects\tabularnewline
Multi-atlas + weighted fusion&20&0.898 (CN) \newline 0.798 (left HC, MCI) &\cite{Wang2011}  &10 rounds of Monte Carlo CV on 20 subjects, pool of 139 (CN/MCI)\tabularnewline
Multi-atlas (MAPS)           &55&0.890                   &\cite{Leung2010}&Segmentation of 30 subjects (10 AD, MCI, and CN)\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: prep
% Prepares the data in a the form needed for plotting.
<<ADNI1-xval-load,include=FALSE,cache=TRUE>>=
jens_atlases = read.csv('data/jens_atlases.csv')
jens_xval_data  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data.csv.gz'))
jens_xval_mean  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data_mean.csv.gz'))
jens_xval_mean$reg_method = factor(jens_xval_mean$reg_method, c("ANTS", "ANIMAL"))
jens_xval_mean$method.mb = factor(jens_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
mb.kappa.best = round(mean(subset(jens_xval_mean, 
                                 reg_method   == "ANTS" & 
                                 method.mb    == "Majority Vote" &
                                 templates.mb == 19 & 
                                 atlases      == 9)$k.mb), digits=3)
mb.diff.best = mean(subset(jens_xval_mean, 
                                 reg_method   == "ANTS" & 
                                 method.mb    == "Majority Vote" &
                                 templates.mb == 19 & 
                                 atlases      == 1)$k_diff)
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

snt.mb.kappa.best =round(mean(subset(snt_xval_mean, 
                                 reg_method   == "ANTS" & 
                                 method.mb    == "Majority Vote" &
                                 templates.mb == 19 & 
                                 atlases      == 9)$k.mb), digits=3)
snt.mb.diff.best = mean(subset(snt_xval_mean, 
                                 reg_method   == "ANTS" & 
                                 method.mb    == "Majority Vote" &
                                 templates.mb == 19 & 
                                 atlases      == 1)$k_diff)
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{color, url, amsmath, geometry, ctable}
\usepackage{titlesec, siunitx, graphicx}
\usepackage[round,authoryear]{natbib}
\usepackage[section]{placeins}   % keep floats in their place.
\usepackage{authblk}             % Listing Author affiliations
\usepackage{rotating}            % rotating figures
\usepackage[hypcap]{caption}
\usepackage{lineno}              % line numbers
\usepackage{multirow,rotating}   % vertical text in tables
\usepackage[displaymath, tightpage]{preview} %draft mode
\setlength{\marginparwidth}{1.2in}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[RO,RE]{draft - December, 2013}

@@

\textbf{Introduction:} 
Advances in image segmentation of magnetic resonance images (MRI) have
demonstrated that \multiatlas{} approaches improve segmentation accuracy and
precision over regular atlas-based approaches.  These approaches often rely on
a large number of such manually segmented atlases (e.g.  30-80) that take
significant time and expertise to produce. We present an algorithm, \mb{} 
({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
T}emplates), for the automatic segmentation of the hippocampus that minimizes
the number of atlases needed while still achieving similar accuracy to
\multiatlas{} approaches. Thus, our method acts as an accurate \multiatlas{}
approach when using special, hard-to-define atlases that are laborious to
construct. \\
\textbf{Method:} \mb{} works by propagating atlas segmentations to a template
library, formed from a subset of target images, via transformations estimated by
nonlinear image registration. The resultant segmentations are then propagated
to each target image and fused using a label fusion method.  To explore the
effect of atlases and templates library size, registation and label
fusion method, we conduct a 10-fold Monte Carlo cross-validation of \mb{} of whole
hippocampal segmentation on 60 ADNI subjects, manually segmented using the
\prsnr{}-protocol, over a range of parameter settings. Using the best settings
found, we then conduct a leave-one-out cross-validation (LOOCV) of hippocampal
subfield segmentation using five high-resolution manually segmented atlases.
Two final experiments assess \mb{} when applied to first episode psychosis and
Alzheimer's disease populations, and \mb{} segmentations are compared with
existing automated methods (\fsl{}, \freesurfer{}, \maper{}) and biases are
explored.\\
\textbf{Results:} Using 9 atlases and 19 template images, \mb{} achieves a
mean Dice's Similarity Coefficient (DSC) of \Sexpr{mb.kappa.best} 
(to \psegs{}) over 10-folds of Monte Carlo % TODO: update cross-validation on 60 subjects, and shows
significantly lower variability in DSC than \multiatlas{} segmentation.
In a LOOCV, \mb{} reproduces the volumes of the cornu ammonis (CA) 1;
CA4/dentate % TODO: update
gyrus (DG); and strata radiatum (SR), strata lacunosum (SL), and strata
moleculare (SM) hippocampal subfields with a percent error in volume that is
at or lower than that produced by image resampling. \mb{} produces hippocampal
volumes in a first episode psychosis patient population that are highly
correlated with expert manual segmentation volumes (Pearson $r = 0.877, t =
16.244, p < 0.001$). Compared to \fsl{} and \freesurfer{}, \mb{} shows much smaller
fixed volume bias (within $250 mm^3$ on average) to semi-automated (\snt{})
segmentations available from ADNI, as well as a conservative, rather than
exaggerated, proportional volume bias.\\ \textbf{Conclusion:} We demonstrate
that \mb{} produces accurate hippocampal segmentations using only 5 atlases over
different hippocampal definitions, disease populations, and acquisition types,
as well as showing that accurate identification of the hippocampal subfields is
possible.
\end{abstract}

@@

We find that for \mb{} segmentations, similarity score increases
as atlas and template library size is increased, although with diminishing returns and
an eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}).  A maximum
similarity score of \Sexpr{mb.kappa.best} is found when using 9
atlases, 19 templates, ANTS registration, and majority vote label fusion).  The
\ants{} registration method consistently out-performs \animal{} registration
over all variable settings we tested. Additionally, by itself, using a weighted
voting strategy did not significantly improve segmentation accuracy, contrary to
the findings of \citet{Aljabar2009} using basic \multiatlas{} segmentation.
Given these findings, in the remainder of this section only results using the
\ants{} registration algorithm and majority vote fusion will be shown.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
\begin{figure}
<<fig:ADNI1-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=7, fig.height=7>>=
stderr <- function(x) sqrt(var(x)/length(x))
ggplot(subset(jens_xval_mean, (templates.mb*atlases) %% 2 == 1 & templates.mb < 21),
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases), linetype=reg_method)) + #as.factor((atlases*templates.mb) %% 2 ))) + 
  facet_grid(.~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
  stat_summary(fun.y=mean,geom='point',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
  stat_summary(fun.y=mean,
               fun.ymin=function (x) mean(x) - stderr(x),
               fun.ymax=function (x) mean(x) + stderr(x), 
               geom='errorbar',
               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases))), linetype='solid', width=0.5) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_y_continuous(breaks=seq(0,1,0.01)) + 
  scale_colour_hue(name="Number of Atlases") +
  scale_linetype(name="Registration Method") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean similarity (DSC)" ) + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text = element_text(size = 8))
@
  \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
  \psegs{} for 60 ADNI subjects vs. atlas and template library
  size, registration algorithm, and label fusion method over 10 folds of Monte Carlo       
  cross validation. 
  Error bars indicate standard error.
  \label{fig:ADNI1-xval-k-mean}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: "increase" in mean kappa over \multiatlas{}
\begin{figure}
<<fig:ADN11-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-load'>>=
ggplot(subset(jens_xval_mean, reg_method=="ANTS" & 
                method.mb=="Majority Vote" & (templates.mb*atlases) %% 2 == 1), 
       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_summary(fun.y=mean, geom="line", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Number of Atlases") +
  scale_alpha_continuous(guide = "none") + 
  scale_linetype_discrete(guide = "none") + 
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean similarity (DSC)" ) +
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{Increase in mean Dice's Similarity Coefficient of \mb{} over
  \multiatlas{} segmentations vs. atlas and template library size when using the
  \ants{} registration method, and majority-vote label fusion over 10 folds of Monte 
  Carlo cross validation. . 
  Segmentation similarity is computed against \psegs{}.  
  Error bars indicate standard error.
  \label{fig:ADNI1-xval-k-diff}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
  #facet_grid( . ~ stat) + 
  geom_line(size=0.5) +
  geom_point(size=2) + 
  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,by=2)) + 
  xlab( "Number of Templates" ) + 
  ylab( "Variability (p)") +
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  scale_size_continuous(guide='none') + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas{} segmentation
  with respect to manual segmentation.}
  Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation
  is computed for each subject across over 10 folds of Monte Carlo       
  cross validation. . 
  Shown on the y-axis is the  p-value resulting from a t-test comparing the
  distribution of variances at each parameter setting (atlas/template library
  size). 
  Only points where \mb{} mean variability is lower than \multiatlas{} are shown.
  Dashed lines indicate a p-value of 0.05 and 0.01.}
  \label{fig:ADNI1-xval-variability}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
\begin{figure}
<<fig:FEP-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=7, fig.height=7>>=
fep.xval <- subset(read.csv('data/results-fep-xval-ants-2013-12-28-020512.csv'),
                   k > 0.5)
fep.xval$atlases <- as.factor(fep.xval$atlases)
fep.xval.mean <- ddply(fep.xval, .(subject,atlases,templates,label), function (df) {
  data.frame(k=mean(df$k),volume=mean(df$volume))
})
stderr <- function(x) sqrt(var(x)/length(x))
ggplot(fep.xval.mean, aes(x=templates,y=k, colour=atlases)) + 
  stat_summary(fun.y=mean,geom='line', aes(y=k,weight=1,group=atlases)) + 
  stat_summary(fun.y=mean,geom='point',aes(y=k,weight=1,group=atlases)) + 
  stat_summary(fun.y=mean,
               fun.ymin=function (x) mean(x) - stderr(x),
               fun.ymax=function (x) mean(x) + stderr(x), 
               geom='errorbar',aes(y=k, weight=1, group=atlases), 
               linetype='solid', width=0.5) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_y_continuous(breaks=seq(0,1,0.01)) + 
  scale_colour_hue(name="Number of Atlases") +
  scale_linetype(name="Registration Method") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean similarity (DSC)" ) + 
  theme(legend.direction = "horizontal", legend.position = "bottom", 
        axis.text = element_text(size = 8))
@
  \caption{Mean Dice's Similarity Coefficient of \mb{} unilateral hippocampal 
  segmentations relative to
  \psegs{} for 81 FEP subjects vs. atlas and template library
  size using \ants registration algorithm and majority vote label fusion method 
  over 3 folds of Monte Carlo cross validation. 
  Error bars indicate standard error.
  \label{fig:FEP-xval-k-mean}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

ggplot(data=fep.volumes, aes(x=manual, y=volume, colour=method)) + 
  geom_smooth(method="lm", formula=y~x) + 
  geom_point() + 
  geom_text(aes(x=3350,y=y,label=eqn),parse=T,data=fep.lm,
      size=3,hjust=0) +
  xlab(expression("Mean manual hippocampus volume " (mm^3))) + 
  ylab(expression("Mean computed hippocampus volume "  (mm^3))) + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{Mean bilateral hippocampus volume as measured by \mb{}, \fsl{}, and
\freesurfer{} vs. manually segmented volumes from the First Episode Patients with
Schizophrenia dataset. \mb{} is run two ways: with an atlas library consisting
of the five Winterburn atlases, or with five manually segmented FEP subjects as
atlases (not included as subjects). Linear fit lines are shown, with shaded
region showing standard error. 
  \label{fig:SZ-volumes}}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tabulation of segmentation accuracies of existing methods (ADNI)
%
%
\begin{sidewaystable}[!tbp]
\scriptsize
\caption{\textbf{Automated segmentation accuracy of the Hippocampus} 
This table shows the best published Dice's overlap measure between automated and
ground truth segmentations of various methods.  Unless otherwise specified,
validation datasets are composed equally of cases and control subjects, and use
manual segmenation labels as ground truth in computing DSC scores. AD =
Alzheimer's Disease; MCI = Mild Cognitive Impairment; CN = Cognitively Normal
(CN); LOOCV = Leave-one-out cross-validation; MCCV = Monte Carlo
cross-validation; SNT = Surgical Medtronic Navigation Technologies
semi-automated labels.  Some studies of automated segmentation of ADNI images
are excluded because they do not not provide overlap measures for the
hippocampus \citep{Heckemann2011,Chupin2009}.
  \label{tab:other-methods}} 
\begin{center}
\begin{tabular}{ p{1.2in}  c  p{1.1in} l  p{3.2in} p{1in} }
\hline\hline
Method & Atlases & DSC \newline {\tiny mean (AD; MCI; CN)} & Reference & Validation & Dataset (Truth)\tabularnewline
\hline
%%  
\mb{}                     &9 & \Sexpr{snt.mb.kappa.best} & \cite{Pipitone2014}
& 10-fold MCCV on 69 subjects 
& ADNI (SNT) 
\tabularnewline 
%%  
Patch-based label fusion  &16&0.861 {\tiny (0.838; ---; 0.883)} &\cite{Coupe2011e}
& LOOCV 
& ADNI (SNT) 
\tabularnewline
%%  
Multi-atlas               &20&0.848 {\tiny (---; 0.798, 0.898)} &\cite{Wang2011}  
& 10-fold MCCV on 20 of 139 subjects
& ADNI (SNT) 
\tabularnewline
%%  
ACM (AdaBoost-based)      &21&0.862                           &\cite{Morra2008} 
& LOOCV                                          
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
LEAP                      &30&0.848                           &\cite{Wolz2010}  
& Segmentation of 182 subjects                   
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
Multi-atlas               &30&0.885                           &\cite{Lotjonen2010}
& Segmentation of 60 subjects                    
& ADNI (SNT)                                           
\tabularnewline                                  
%%                                               
Multi-atlas (MAPS)        &55&0.890                           &\cite{Leung2010}
& Segmentation of 30 subjects (10 AD, MCI, and CN)
& ADNI (SNT) 
\tabularnewline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hline
%& & {\tiny mean (right; left)} & & & \tabularnewline
%\hline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
\mb{}                      &9 & \Sexpr{mb.kappa.best}       &\cite{Pipitone2014} 
& 10-fold MCCV on 60 subjects 
& ADNI (\prsnr)
\tabularnewline 
%%
Neural nets                &10&0.740   &\cite{Powell2008} % {\tiny (0.72, 0.75)} 
& Segmentation of 5 subjects 
& controls %Other
\tabularnewline
%%
Probablistic atlas         &11&0.852   &\cite{VanderLijn2008}
& 11 atlases used in 100 rounds of LOOCV on 20 elderly subjects
& elderly controls %Other
\tabularnewline
%%
Probabilistic Atlas        &16&0.860   &\cite{Chupin2009}
& LOOCV 
& AD subjects %Other
\tabularnewline
%%
Anatomically-guided EM     &17&0.812   &\cite{Pohl2007} %{\tiny (0.815, 0.808)} 
& LOOCV on 17 controls, segmentation of 33 mixed subjects
& mixed diagnosis %Other
\tabularnewline
%%
Multi-atlas                &30&0.820   &\cite{Heckemann2006} %{\tiny (0.83, 0.81)}    
& LOOCV
& controls %Other
\tabularnewline
%%
Multi-atlas                &30&0.880   &\cite{Gousias2008} %{\tiny (0.88,---)}    
& 30 adult atlas used, segmentation of 33 2yr old subjects 
& 2yr old controls  %Other
\tabularnewline
%%
Multi-atlas                &80&0.890   &\cite{Collins2010}
& LOOCV                 
& controls %Other       
\tabularnewline
%%                      
Multi-atlas                &55&0.860   &\cite{Barnes2008}%{\tiny (---,0.86)}    
& LOOCV 
& controls and AD %Other
\tabularnewline
%%
Multi-atlas               &275&0.835   &\cite{Aljabar2009}% {\tiny (0.832, 0.837)} 
& LOOCV 
& controls %Other
\tabularnewline
\hline
\end{tabular}
\end{center}
\end{sidewaystable}
